---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(tidyverse)
```

# blblm

The goal of blblm is to apply the Bag of Little Bootstraps method to Linear Regression. This package draws inspiration from `stats::lm()` and its related functions. It also incorporates `Rcpp` and `furrr` to improve performance as well as some basic `Tidyverse` principles to improve the elegance of the code. 

For full documentation, please see the vignette.

## Installation

You can install `blblm` from here using:

``` r
devtools::devtools::install_github("a10thehero/blblm")
```

## Example

The vignette should be consulted to get a better understanding of the package, but here are some usage examples anyways.

```{r example}
library(blblm)


future::plan(future::multiprocess, workers = 2)
mtcars %>% blblm(mpg ~ cyl, m = 2, B = 500, useParallel = TRUE) %>% Adj.Rsq.blblm(confidence = TRUE, level = 0.95)


mtcars %>% blblm(mpg ~ 0 + hp * cyl, m = 2, B = 101) %>% coef()


mtcars %>% blblm(wt ~ mpg, m = 3, B = 200) %>% confint(level = 0.99)
```